---
title: Hatching and Settling Variance Components
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
source("../0_misc_funcs.R")
load("Model_III_posterior_20250930_1554.rdata")
```

## Run Summary  

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Diagnostics
```{r include=FALSE}
smry <- summary(
  post, 
  vars = c("deviance", "sire.vcov", "dam.vcov", "interaction.vcov")
) 
```
```{r}
smry |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  

## V~A~ on the observed scale   
```{r}
# p$VA is latent scale variance
comp.smry(vcv.obs$VA$vcv.G.obs)
plot.metric(vcv.obs$VA$vcv.G.obs)
```

## V~M~ on the observed scale   
```{r}
comp.smry(vcv.obs$VM$vcv.G.obs)
plot.metric(vcv.obs$VM$vcv.G.obs)
```
\newpage  

## V~D~ on the observed scale   
```{r}
comp.smry(vcv.obs$VD$vcv.G.obs)
plot.metric(vcv.obs$VD$vcv.G.obs)
```
\newpage  

## V~M~ / V~A~ on the observed scale   
```{r}
vm.va <- vcv.obs$VM$vcv.G.obs / vcv.obs$VA$vcv.G.obs
comp.smry(vm.va, 1)
plot.metric(vm.va) + scale_x_log10()
```

## V~D~ / V~A~ on the observed scale   
```{r}
vd.va <- vcv.obs$VD$vcv.G.obs / vcv.obs$VA$vcv.G.obs
comp.smry(vd.va, 1)
plot.metric(vd.va) + scale_x_log10()
```
\newpage  

## Heritability on the observed scale   
```{r}
comp.smry(p$H)

p$H |>
  t() |> 
  as.data.frame() |>
  pivot_longer(everything()) |>
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Value", y = "Count") +
  facet_wrap(~ name, scales = "free")
```
\newpage  

## Evolvability (single traits) on the observed scale   
```{r}
comp.smry(p$E)

p$E |>
  t() |> 
  as.data.frame() |>
  pivot_longer(everything()) |>
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Value", y = "Count") +
  facet_wrap(~ name, scales = "free")
```

## Settling / Hatching Evolvability
```{r}
e.ratio <- p$E[2, ] / p$E[1, ]

e.ratio |> 
  log() |> 
  vecSmry() |>
  exp() |> 
  rbind() |>
  as.data.frame() |>
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)

data.frame(ratio = e.ratio) |> 
  ggplot(aes(ratio)) +
  geom_histogram(bins = 100) +
  scale_x_log10() +
  labs(x = "Settling / Hatching Evolvability", y = "Count")
```


## Evolvability (from G matrix)    
This is for a random set of selection gradients.  
```{r}
# Get the posterior samples by averaging across all of the random beta's
eB_posterior_hatch_settle <- apply(e.params_BetaMCMC$post.dist$eB, 1, mean)
rB_posterior_hatch_settle <- apply(e.params_BetaMCMC$post.dist$rB, 1, mean)
cB_posterior_hatch_settle <- apply(e.params_BetaMCMC$post.dist$cB, 1, mean)
aB_posterior_hatch_settle <- apply(e.params_BetaMCMC$post.dist$aB, 1, mean)
iB_posterior_hatch_settle <- apply(e.params_BetaMCMC$post.dist$iB, 1, mean)

E_hatch_settle <- 
  data.frame(name = c(rep("eB",length(eB_posterior_hatch_settle)),
                      rep("rB",length(rB_posterior_hatch_settle)),
                      rep("cB",length(cB_posterior_hatch_settle)),
                      rep("aB",length(aB_posterior_hatch_settle)),
                      rep("iB",length(iB_posterior_hatch_settle))),
             value = c(eB_posterior_hatch_settle,
                       rB_posterior_hatch_settle,
                       cB_posterior_hatch_settle,
                       aB_posterior_hatch_settle,
                       iB_posterior_hatch_settle))

E_hatch_settle <- E_hatch_settle |> 
  mutate(name = factor(name, levels = c("eB","rB","cB", "iB", "aB"))) |>
  group_by(name) |>
  summarise(summary = list(round(vecSmry(value), 6)), .groups = "drop") |> unnest_wider(summary)

pander::pander(E_hatch_settle)
```
