---
title: Summary of Model IV
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
load("Model_IV_posterior_20251001_0640.rdata")
source("../0_misc_funcs.R")
```

## Run Summary

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Summary
```{r, include=FALSE}
smry <- summary(
  post, 
  vars = c("sire.vcov", "dam.vcov", 
    "interaction.vcov", "resid.vcov"
  )
)
```
```{r}
smry |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  

## G matrix  
```{r}
G_matrix <- p$VA[1:2,1:2,]
comp.smry(G_matrix)
plot.metric(G_matrix)
```


## Selection differentials, s<sub>g</sub> = cov<sub>g</sub>(w, z)  
- We have Cov(length, p) on the data scale, from QGmvparams() 
- We want Cov(length, w) where w = p / mean(p) (relative fitness)  
- Use linear scaling property of covariance, so that:  
Cov(length, w) = Cov(length, p / mean(p)) = 1 / mean(p) * Cov(length, p)  
- So we just divide Cov(length, p) by the mean of p  

Trunk selection differential  
```{r}
cov_trunk_p <- vcv.obs$vcv.G.obs[1,3,]
mean_p <- vcv.obs$mean.obs[, "Settling"]
cov_trunk_p_relative <- cov_trunk_p / mean_p # selection differential
vecSmry(cov_trunk_p_relative)
plot.metric(cov_trunk_p_relative)
```

Tail selection differential
```{r}
cov_tail_p <- vcv.obs$vcv.G.obs[2,3,]
mean_p <- vcv.obs$mean.obs[, "Settling"]
cov_tail_p_relative <- cov_tail_p / mean_p # selection differential
vecSmry(cov_tail_p_relative)
plot.metric(cov_tail_p_relative)
```

## Selection gradients, &beta;<sub>g</sub> = G<sup>-1</sup>s<sub>g</sub>  
```{r}
sg <- rbind(cov_trunk_p_relative, cov_tail_p_relative)  # 2 x n_iter

n_iter <- dim(G_matrix[,,])[3]

beta <- matrix(NA, nrow = 2, ncol = n_iter,
                     dimnames = list(c("Trunk", "Tail"), NULL))

for (i in seq_len(n_iter)) {
  G <- G_matrix[, , i]
  S_vec <- sg[, i]
  beta[, i] <- solve(G) %*% S_vec
}

comp.smry(beta)
plot.metric(beta[1,])
plot.metric(beta[2,])
```

## Evolvability  
Predicts the % change in trunk and tail length per generation (assuming fitness is settlement)   
```{r}
# Extract means for convenience
mean_trunk <- p$mean.overall[1,]
mean_tail  <- p$mean.overall[2,]

# Extract variances and covariance
VA11 <- p$VA[1,1,]  # Var(Trunk)
VA22 <- p$VA[2,2,]  # Var(Tail)
VA12 <- p$VA[1,2,]  # Cov(Trunk, Tail)

# Compute trunk and tail evolvability
E_matrix <- rbind(
  Trunk = (VA11 / mean_trunk^2) * beta[1, ] + (VA12 / (mean_trunk * mean_tail)) * beta[2, ],
  Tail  = (VA22 / mean_tail^2)  * beta[2, ] + (VA12 / (mean_trunk * mean_tail)) * beta[1, ]
)

comp.smry(E_matrix * 100)
```

Trunk  
```{r}
plot.metric(E_matrix['Trunk',] * 100) # needs more bins in geom_histogram()
```

Tail  
```{r}
plot.metric(E_matrix['Tail',] * 100) # needs more bins in geom_histogram()
```

Tail evolvability / Trunk evolvability
```{r}
vecSmry(E_matrix['Tail',] / E_matrix['Trunk',])
plot.metric(E_matrix['Tail',] / E_matrix['Trunk',]) # needs more bins in geom_histogram()
```