---
title: Summary of Model I (Trunk, Tail)
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
load("Model_I_posterior_20250930_0019.rdata")
source("../0_misc_funcs.R")
```

## Run Summary

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Summary
```{r, include=FALSE}
smry <- summary(
  post, 
  vars = c(
    "deviance", "sire.vcov", "dam.vcov", 
    "interaction.vcov", "resid.vcov"
  )
)
```
```{r}
smry |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  

## V~A~
```{r}
comp.smry(p$VA)
plot.metric(p$VA)
```

## V~M~
```{r}
comp.smry(p$VM)
plot.metric(p$VM)
```
\newpage  

## V~D~
```{r}
comp.smry(p$VD)
plot.metric(p$VD)
```

## V~P~
```{r}
comp.smry(p$VP)
plot.metric(p$VP)
```
\newpage  

## V~M~ / V~A~
```{r}
vm.va <- p$VM / p$VA
comp.smry(vm.va, 1)
plot.metric(vm.va) + scale_x_log10()
```

## V~D~ / V~A~
```{r}
vd.va <- p$VD / p$VA
comp.smry(vd.va, 1)
plot.metric(vd.va) + scale_x_log10()
```
\newpage  

## Heritability
```{r}
comp.smry(p$H)
plot.metric(p$H)
```
\newpage  

## Evolvability (single traits)
```{r}
comp.smry(p$E)

p$E |> 
  t() |> 
  as.data.frame() |> 
  pivot_longer(everything()) |> 
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Value", y = "Count") +
  facet_wrap(~ name, scales = "free")
```

## Tail / Trunk Evolvability
```{r}
e.ratio <- p$E[2, ] / p$E[1, ]

e.ratio |> 
  log() |> 
  vecSmry() |>
  exp() |> 
  rbind() |>
  as.data.frame() |>
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)

data.frame(ratio = e.ratio) |> 
  ggplot(aes(ratio)) +
  geom_histogram(bins = 100) +
  scale_x_log10() +
  labs(x = "Tail / Trunk Evolvability", y = "Count")
```

## Evolvability (from G matrix) v1  
This is for a random set of selection gradients.
```{r}
# Get the posterior samples by averaging across all of the random beta's
eB_posterior_head_tail <- apply(e.params_BetaMCMC$post.dist$eB, 1, mean)
rB_posterior_head_tail <- apply(e.params_BetaMCMC$post.dist$rB, 1, mean)
cB_posterior_head_tail <- apply(e.params_BetaMCMC$post.dist$cB, 1, mean)
aB_posterior_head_tail <- apply(e.params_BetaMCMC$post.dist$aB, 1, mean)
iB_posterior_head_tail <- apply(e.params_BetaMCMC$post.dist$iB, 1, mean)

E_trunk_tail <- 
  data.frame(name = c(rep("eB",length(eB_posterior_head_tail)),
                      rep("rB",length(rB_posterior_head_tail)),
                      rep("cB",length(cB_posterior_head_tail)),
                      rep("aB",length(aB_posterior_head_tail)),
                      rep("iB",length(iB_posterior_head_tail))),
             value = c(eB_posterior_head_tail,
                       rB_posterior_head_tail,
                       cB_posterior_head_tail,
                       aB_posterior_head_tail,
                       iB_posterior_head_tail))

E_trunk_tail <- E_trunk_tail |> 
  mutate(name = factor(name, levels = c("eB","rB","cB", "iB", "aB"))) |>
  group_by(name) |>
  summarise(summary = list(round(vecSmry(value), 6)), .groups = "drop") |> unnest_wider(summary)

pander::pander(E_trunk_tail)
```

## Evolvability (from G matrix) v2 
This is for a specific set of selection gradients:    
Beta_index = 1 is $\beta = \begin{bmatrix} 0 \\ 1 \end{bmatrix}$ = strong selection for longer tails only  
Beta_index = 2 is $\beta = \begin{bmatrix} -1 \\ -1 \end{bmatrix}$ = strong selection for shorter trunks and shorter tails  
Beta_index = 3 is $\beta = \begin{bmatrix} 1 \\ -1 \end{bmatrix}$ = strong selection for longer trunks and shorter tails  

```{r}
summary_stacked <- e.params_beta %>%
  pivot_longer(cols = c(e, r, c, a , i), 
               names_to = "metric", 
               values_to = "value") %>%
    mutate(metric = recode(metric,
                         e = "evolvability",
                         r = "respondability",
                         c = "conditional",
                         a = "autonomy (cB/eB)",
                         i = "integration"),
         metric = factor(metric, 
                         levels = c("evolvability",
                                    "respondability",
                                    "conditional",
                                    "autonomy (cB/eB)",
                                    "integration"))) %>% 
  group_by(Beta_index, metric) %>%
  summarise(summary = list(round(vecSmry(value), 6)), .groups = "drop") %>%
  unnest_wider(summary) %>%
  arrange(metric, Beta_index)

pander::pander(summary_stacked)
```
