---
title: Summary of Model II
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
source("../0_misc_funcs.R")
load("Model_II_posterior_20250921_0942.rdata")
```

## Run Summary

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Summary
```{r, include=FALSE}
smry <- summary(
  post, 
  vars = c(
    "deviance", "sire.var", "dam.var", "interaction.var", "resid.var"
  )
)
```
```{r}
smry |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  


## V~A~ on the observed scale   
```{r}
# p$VA is latent scale variance
comp.smry(var.obs$VA$var.a.obs)
plot.metric(var.obs$VA$var.a.obs)
```

## V~M~ on the observed scale    
```{r}
# p$VM is latent scale variance
comp.smry(var.obs$VM$var.a.obs)
plot.metric(var.obs$VM$var.a.obs)
```
\newpage  

## V~D~ on the observed scale   
```{r}
# p$VD is latent scale variance
comp.smry(var.obs$VD$var.a.obs)
plot.metric(var.obs$VD$var.a.obs)
```

## V~P~ on the observed scale    
```{r}
# p$VP is latent scale variance
comp.smry(var.obs$VA$var.obs)
plot.metric(var.obs$VA$var.obs)
```
\newpage  

## V~M~ / V~A~ on the observed scale    
```{r}
vm.va <- var.obs$VM$var.a.obs / var.obs$VA$var.a.obs
comp.smry(vm.va, 1)
plot.metric(vm.va) + scale_x_log10()
```

## V~D~ / V~A~ on the observed scale    
```{r}
vd.va <- var.obs$VD$var.a.obs / var.obs$VA$var.a.obs
comp.smry(vd.va, 1)
plot.metric(vd.va) + scale_x_log10()
```
\newpage  

## Heritability on the observed scale    
```{r}
comp.smry(p$H)
plot.metric(p$H)
```
\newpage  

## Evolvability on the observed scale   
```{r}
comp.smry(p$E)

p$E |> 
  enframe() |> 
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Value", y = "Count") 
```
