---
title: Summary of Model IV
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
load("Model_IV_posterior_20250918_0443.rdata")
source("../0_misc_funcs.R")
```

## Run Summary

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Summary
```{r, include=FALSE}
smry <- summary(
  post, 
  vars = c("sire.vcov", "dam.vcov", 
    "interaction.vcov", "resid.vcov"
  )
)
```
```{r}
smry |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  

## G matrix  
```{r}
comp.smry(p$VA[1:2,1:2,])
plot.metric(p$VA[1:2,1:2,])
```


## Selection differentials, s<sub>g</sub> = cov<sub>g</sub>(w, z)  
Notes: 1) shown here are on the length x logit scale, we need on length x probability scale. QGglmm::QGparams needs to evaluate the whole matrix, not just the diagonals. 2) we need relative fitess = p / mean(p)...I think Var(p/mean(p)) = Var(p) / mean(p)^2, so var.obs / mean.obs^2  ?  
Trunk - Settling. 
```{r}
vecSmry(p$VA[3,1,])
```

Tail - Settling
```{r}
vecSmry(p$VA[3,2,])
```

## Selection gradients, &beta;<sub>g</sub> = G<sup>-1</sup>s<sub>g</sub>  
Work in progress...  
```{r}
comp.smry(p$VA[1:2,1:2,])
```
