---
title: Hatching and Settling Variance Components
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
source("0_misc_funcs.R")

load("hatch_settle_posterior_20230119_1741.rdata")
```

## Run Summary  

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Diagnostics
```{r include=FALSE}
smry <- summary(
  post, 
  vars = c("deviance", "additive.vcov", "maternal.vcov", "interaction.vcov")
) 
```
```{r}
smry %>% 
  as.data.frame() %>% 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  

## V~A~
```{r}
comp.smry(p$VA)
plot.metric(p$VA)
```

## V~M~
```{r}
comp.smry(p$VM)
plot.metric(p$VM)
```
\newpage  

## V~D~
```{r}
comp.smry(p$VD)
plot.metric(p$VD)
```
\newpage  

## V~M~ / V~A~
```{r}
vm.va <- p$VM / p$VA
comp.smry(vm.va, 1)
plot.metric(vm.va) + scale_x_log10()
```

## V~D~ / V~A~
```{r}
vd.va <- p$VD / p$VA
comp.smry(vd.va, 1)
plot.metric(vd.va) + scale_x_log10()
```
\newpage  

## Heritability
```{r}
comp.smry(p$H)

p$H %>%
  t() %>% 
  as.data.frame() %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Value", y = "Count") +
  facet_wrap(~ name, scales = "free")
```
\newpage  

## Evolvability
```{r}
comp.smry(p$E)

p$E %>%
  t() %>% 
  as.data.frame() %>%
  pivot_longer(everything()) %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Value", y = "Count") +
  facet_wrap(~ name, scales = "free")
```

## Settling / Hatching Evolvability
```{r}
e.ratio <- p$E[2, ] / p$E[1, ]

e.ratio %>% 
  log() %>% 
  vecSmry() %>%
  exp() %>% 
  rbind() %>%
  as.data.frame() %>%
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)

data.frame(ratio = e.ratio) %>% 
  ggplot(aes(ratio)) +
  geom_histogram(bins = 100) +
  scale_x_log10() +
  labs(x = "Settling / Hatching Evolvability", y = "Count")
```