---
title: Settling Prediction
output:
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
load("Model_IV_posterior_20230117_0038.rdata")
source("0_misc_funcs.R")
```

## Run Summary

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Summary
```{r, include=FALSE}
smry <- summary(
  post, 
  vars = c(
    "deviance", "intercept", "int.beta", "sire.beta", "maternal.beta", "block.beta"
  )
)
```
```{r}
smry %>% 
  as.data.frame() %>% 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```

## Interaction
```{r, fig.width = 8, fig.height = 8}
n.pts <- 1000
pr.breaks <- seq(0, 1, length.out = n.pts)

for(m in dimnames(p$interaction.mean)[[2]]) {
  int.breaks <- seq(
    floor(min(p$interaction.mean[, m, ])), 
    ceiling(max(p$interaction.mean[, m, ])),
    length.out = n.pts
  )
  
  pr.settle <- parallel::mclapply(int.breaks, function(x) {
    sapply(1:model.data$n.interactions, function(i) {
      p$intercept[m, ] + 
        (p$int.beta[m, ] * x) +
        (p$sire.beta[m, ] * t(p$additive.sire.eff[model.data$sire[i], m, ])) +
        (p$maternal.beta[m, ] * t(p$maternal.eff[model.data$dam[i], m, ])) +
        (p$block.beta[m, ] * t(p$block.eff[model.data$block[i], m, ]))
    }) %>%
      as.vector() %>%
      plogis() %>%
      cut(pr.breaks, include.lowest = TRUE) %>%
      table()
  }, mc.cores = 10) %>% 
    do.call(cbind, .) %>% 
    as.data.frame() %>% 
    remove_rownames() %>% 
    setNames(int.breaks) %>% 
    mutate(pr.settle = apply(cbind(pr.breaks[-n.pts], pr.breaks[-1]), 1, mean)) %>% 
    pivot_longer(-pr.settle, names_to = "interaction.mean", values_to = "freq") %>% 
    mutate(
      interaction.mean = as.numeric(interaction.mean),
      interaction.mean.lik = dnorm(interaction.mean, mean(p$interaction.mean[, m, ]), sd(p$interaction.mean[, m, ])),
      wt = freq * interaction.mean.lik
    )
  
  gg <- ggplot(pr.settle, aes(interaction.mean, pr.settle)) +
    geom_tile(aes(fill = wt)) +
    geom_hline(yintercept = 0.5, color = "white", alpha = 0.6, linetype = "dashed", linewidth = 1) +
    scale_fill_viridis_c(option = "inferno") +
    labs(x = paste("mean interaction", m, "length"), y = "Pr(settle)") +
    coord_cartesian(xlim = range(int.breaks), ylim = c(0, 1), expand = FALSE) +
    theme_minimal() +
    theme(
      legend.position = "none",
      panel.grid = element_blank()
    )
  plot(gg)
}
```
  
### Beta summary
```{r}
pct.gt.0 <- apply(p$int.beta, 1, function(x) mean(x > 0))
p$int.beta %>% 
  apply(1, vecSmry) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(pct.gt.0 = pct.gt.0) %>% 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
### Effect summary
```{r, fig.width = 12, fig.height = 30}
as.data.frame.table(p$interaction.mean) %>% 
  setNames(c("int", "metric", "iter", "effect")) %>% 
  mutate(
    int = as.numeric(int),
    iter = as.numeric(as.character(iter))
  ) %>% 
  ggplot(aes(effect)) +
  geom_histogram(bins = 50, alpha = 0.7) +
  labs(x = "Interaction mean", y = "Count") +
  facet_grid(int ~ metric, scales = "free_x")
```

## Marginal Effects
```{r, fig.width = 8, fig.height = 8}
pred.pr.settle %>% 
  filter(effect.type != "Interaction Mean") %>% 
  ggplot(aes(x = effect)) + 
  geom_hline(yintercept = mean(model.data$settle[, 1]), linetype = "dashed") +
  geom_line(aes(y = median, color = effect.type)) +
  labs(x = "Effect", y = "Pr(settling)") +
  facet_wrap(~ metric) +
  theme(legend.position = "top")
```

## Maternal
### Beta summary
```{r}
pct.gt.0 <- apply(p$maternal.beta, 1, function(x) mean(x > 0))
p$maternal.beta %>% 
  apply(1, vecSmry) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(pct.gt.0 = pct.gt.0) %>% 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
### Effect summary
```{r, fig.width = 12, fig.height = 12}
as.data.frame.table(p$maternal.eff) %>% 
  setNames(c("dam", "metric", "iter", "effect")) %>% 
  mutate(
    dam = as.numeric(dam),
    iter = as.numeric(as.character(iter))
  ) %>% 
  ggplot(aes(effect)) +
  geom_vline(xintercept = 0, color = "red") +
  geom_histogram(bins = 50, alpha = 0.7) +
  xlim(
    pred.pr.settle %>% 
      filter(effect.type == "Maternal") %>% 
      pull(effect) %>% 
      range()
  ) +
  labs(x = "Maternal effect", y = "Count") +
  facet_grid(dam ~ metric)
```

## Additive Sire
### Beta summary
```{r}
pct.gt.0 <- apply(p$sire.beta, 1, function(x) mean(x > 0))
p$sire.beta %>% 
  apply(1, vecSmry) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(pct.gt.0 = pct.gt.0) %>% 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
### Effect summary
```{r, fig.width = 12, fig.height = 12}
as.data.frame.table(p$additive.sire.eff) %>% 
  setNames(c("sire", "metric", "iter", "effect")) %>% 
  mutate(
    sire = as.numeric(sire),
    iter = as.numeric(as.character(iter))
  ) %>% 
  ggplot(aes(effect)) +
  geom_vline(xintercept = 0, color = "red") +
  geom_histogram(bins = 50, alpha = 0.7) +
  xlim(
    pred.pr.settle %>% 
      filter(effect.type == "Additive Sire") %>% 
      pull(effect) %>% 
      range()
  ) +
  labs(x = "Additive sire effect", y = "Count") +
  facet_grid(sire ~ metric)
```

## Block
### Beta summary
```{r}
pct.gt.0 <- apply(p$block.beta, 1, function(x) mean(x > 0))
p$block.beta %>% 
  apply(1, vecSmry) %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate(pct.gt.0 = pct.gt.0) %>% 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
### Effect summary
```{r, fig.width = 12, fig.height = 12}
as.data.frame.table(p$block.eff) %>% 
  setNames(c("block", "metric", "iter", "effect")) %>% 
  mutate(
    block = as.numeric(block),
    iter = as.numeric(as.character(iter))
  ) %>% 
  ggplot(aes(effect)) +
  geom_vline(xintercept = 0, color = "red") +
  geom_histogram(bins = 50, alpha = 0.7) +
  xlim(
    pred.pr.settle %>% 
      filter(effect.type == "Block") %>% 
      pull(effect) %>% 
      range()
  ) +
  labs(x = "Block effect", y = "Count") +
  facet_grid(block ~ metric)
```