---
title: Summary of Model IV
output:
  html_document:
    toc: yes
    toc_float: yes
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
                      encoding = encoding,
                      output_dir = "Figures and Tables")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
source("0_misc_funcs.R")
load("Model_outputs/Model_IV_posterior_20251001_0640.rdata")
```

## Run Summary

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Summary
```{r, include=FALSE}
smry <- summary(
  post, 
  vars = c("sire.vcov", "dam.vcov", 
    "interaction.vcov", "resid.vcov"
  )
)
```
```{r}
smry |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  

## G matrix  
```{r}
G_matrix <- p$VA[1:2,1:2,]
comp.smry(G_matrix)
plot.metric(G_matrix)
```


## Selection differentials, s<sub>g</sub> = cov<sub>g</sub>(w, z)  
- We have Cov(p, length) on the data scale, from QGmvparams() 
- We want Cov(w, length) where w = p / mean(p) (relative fitness)  
- Use linear scaling property of covariance, so that:  
Cov(w, length) = Cov(p / mean(p), length) = 1 / mean(p) * Cov(p, length)  
- So we just divide Cov(p, length) by the mean of p  

Trunk selection differential  
```{r}
cov_trunk_p <- vcv.obs$vcv.G.obs[1,3,]
mean_p <- vcv.obs$mean.obs[, "Settling"]
cov_trunk_p_relative <- cov_trunk_p / mean_p # selection differential
c(
  vecSmry(cov_trunk_p_relative),
  pr.gt.0 = mean(cov_trunk_p_relative > 0)
)
plot.metric(cov_trunk_p_relative)
```

Tail selection differential
```{r}
cov_tail_p <- vcv.obs$vcv.G.obs[2,3,]
mean_p <- vcv.obs$mean.obs[, "Settling"]
cov_tail_p_relative <- cov_tail_p / mean_p # selection differential
c(
  vecSmry(cov_tail_p_relative),
  pr.gt.0 = mean(cov_tail_p_relative > 0)
)
plot.metric(cov_tail_p_relative)
```

## Selection gradients, &beta;<sub>g</sub> = G<sup>-1</sup>s<sub>g</sub>  
```{r}
sg <- rbind(cov_trunk_p_relative, cov_tail_p_relative)  # 2 x n_iter

n_iter <- dim(G_matrix[,,])[3]

beta <- matrix(NA, nrow = 2, ncol = n_iter,
                     dimnames = list(c("Trunk", "Tail"), NULL))

for (i in seq_len(n_iter)) {
  G <- G_matrix[, , i]
  S_vec <- sg[, i]
  beta[, i] <- solve(G) %*% S_vec
}

beta |> 
  apply(1, function(x) {
    c(vecSmry(x), pr.gt.0 = mean(x > 0))
  }) |> 
  t() |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
plot.metric(beta[1,])
plot.metric(beta[2,])
```

## Evolvability  
Predicts the % change in trunk and tail length per generation (assuming fitness is settlement)   
\[
\frac{\Delta z_i}{\bar{z}_i} \times 100
=
\frac{s_{g,i}}{\bar{z}_i} \times 100
\qquad \text{for each trait } i.
\]
```{r}
# Extract means
mean_z <- rbind(p$mean.overall[1,], p$mean.overall[2,])

# Compute trunk and tail evolvability
n_iter <- dim(sg)[2]

E_matrix <- matrix(NA, nrow = 2, ncol = n_iter,
                     dimnames = list(c("Trunk", "Tail"), NULL))

for (i in seq_len(n_iter)) {
  S_vec <- sg[, i]
  mean_vec <- mean_z[, i]
  E_matrix[, i] <- S_vec / mean_vec
}

(E_matrix * 100) |> 
  apply(1, function(x) {
    c(vecSmry(x), pr.gt.0 = mean(x > 0))
  }) |> 
  t() |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```

Trunk  
```{r}
(E_matrix['Trunk', ] * 100) |> 
  enframe() |>
  ggplot(aes(value)) +
  geom_histogram(bins = 200) +
  labs(x = "Trunk Evolvability", y = "Count")
```

Tail  
```{r}
(E_matrix['Tail', ] * 100) |> 
  enframe() |>
  ggplot(aes(value)) +
  geom_histogram(bins = 200) +
  labs(x = "Tail Evolvability", y = "Count")# needs more bins in geom_histogram()
```

Tail evolvability / Trunk evolvability
```{r}
e.ratio <- E_matrix['Tail', ] / E_matrix['Trunk', ]
c(vecSmry(e.ratio), pr.gt.1 = mean(e.ratio > 1))
e.ratio |> 
  enframe() |>
  ggplot(aes(value)) +
  geom_histogram(bins = 200) +
  labs(x = "Tail Evolvability / Trunk Evolvability", y = "Count")
```

