---
title: Summary of JAGS model
output:
  html_document:
    toc: yes
    toc_float: yes
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
                      encoding = encoding,
                      output_dir = "Figures and Tables")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 3,
  fig.width = 8
)
options(scipen = 10, digits = 3)

rm(list = ls())
library(tidyverse)
library(runjags)
load("Model_outputs/Model_II_posterior_20260605_0120.rdata")
source("0_misc_funcs.R")
elapsed <- swfscMisc::autoUnits(post$timetaken)
```

## Run Summary

Chains : `r length(post$mcmc)`  
Burn-in : `r post$burnin`  
Samples per chain : `r post$sample`  
Total samples : `r post$sample * length(post$mcmc)`  
Thinning : `r post$thin`   
Duration: `r paste(round(elapsed, 1), units(elapsed))`  
Start Time: `r format(start, "%Y-%m-%d %H:%M")`  
End Time: `r format(end, "%Y-%m-%d %H:%M")`   

## Posterior Summary
```{r, include=FALSE}
smry <- summary(
  post, 
  vars = c("deviance", "sire.vcov", "dam.vcov", "int.vcov", "resid.vcov")
)
```
```{r}
smry |> 
  as.data.frame() |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
\newpage  

## V~A~   
Latent scale  
```{r}
smry <- cov.smry(p$VA)
 pander::pander(smry$smry, split.tables = Inf, keep.line.breaks = TRUE) 
plot(smry$plot)
```

## V~M~  
Latent scale  
```{r}
smry <- cov.smry(p$VM)
pander::pander(smry$smry, split.tables = Inf, keep.line.breaks = TRUE) 
plot(smry$plot)
```
\newpage  

## V~D~  
Latent scale  
```{r}
smry <- cov.smry(p$VD)
 pander::pander(smry$smry, split.tables = Inf, keep.line.breaks = TRUE) 
plot(smry$plot)
```

## V~R~ (latent scale)
```{r}
smry <- cov.smry(p$resid)
 pander::pander(smry$smry, split.tables = Inf, keep.line.breaks = TRUE) 
plot(smry$plot)
```

## V~P~  
Latent scale  
```{r}
smry <- cov.smry(p$VP)
 pander::pander(smry$smry, split.tables = Inf, keep.line.breaks = TRUE) 
plot(smry$plot)
```
\newpage  

## V~M~ / V~A~  
Latent scale  
```{r}
smry <- cov.smry(p$VM / p$VA, 1)
 pander::pander(smry$smry, split.tables = Inf, keep.line.breaks = TRUE) 
plot(smry$plot + scale_x_log10())
```

## V~D~ / V~A~   
Latent scale
```{r}
smry <- cov.smry(p$VD / p$VA, 1)
 pander::pander(smry$smry, split.tables = Inf, keep.line.breaks = TRUE) 
plot(smry$plot + scale_x_log10())
```
\newpage  

## Heritability 
```{r}
smry <- apply(p$H,1,vecSmry)
 pander::pander(smry, split.tables = Inf, keep.line.breaks = TRUE) 

 p$H |>
  t() |>
  as.data.frame() |>
  pivot_longer(everything()) |>
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Value", y = "Count") +
  facet_wrap(~ name, scales = "free")
```
\newpage  

## Evolvability (for single traits)
```{r}
smry <- apply(p$E,1,vecSmry)
 pander::pander(smry, split.tables = Inf, keep.line.breaks = TRUE)
p$E |>
  t() |>
  as.data.frame() |>
  pivot_longer(everything()) |>
  ggplot(aes(value)) +
  geom_histogram(bins = 100) +
  labs(x = "Evolvability", y = "Count") +
  facet_wrap(~ name, scales = "free")
```
Evolvability as a percentage
```{r}
smry <- apply(p$E * 100,1,vecSmry)
 pander::pander(smry, split.tables = Inf, keep.line.breaks = TRUE)
p$E |>
  t() |>
  as.data.frame() |>
  pivot_longer(everything()) |>
  ggplot(aes(value * 100)) +
  geom_histogram(bins = 100) +
  labs(x = "Evolvability (%)", y = "Count") +
  facet_wrap(~ name, scales = "free")
```

## Tail / Trunk Evolvability
```{r}
e.ratio <- p$E[2, ] / p$E[1, ]

e.ratio |>
  log() |>
  vecSmry() |>
  exp() |>
  rbind() |>
  as.data.frame() |>
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)

data.frame(ratio = e.ratio) |>
  ggplot(aes(ratio)) +
  geom_histogram(bins = 100) +
  scale_x_log10() +
  labs(x = "Tail / Trunk Evolvability", y = "Count")
```

## Evolvability (from G matrix) v1  
This is for a random set of selection gradients.  
Multiply eB by 100 to get in units of percent.  
```{r}
# Get the posterior samples by averaging across all of the random beta's
e.names <- c('eB', 'rB', 'cB', 'iB', 'aB')

lapply(e.names, function(x) {
  data.frame(
    value = apply(e.params_BetaMCMC$post.dist[[x]], 1, mean),
    name = x
  )
}) |> 
  bind_rows() |> 
  mutate(name = factor(name, levels = e.names)) |>
  group_by(name) |>
  summarise(summary = list(round(vecSmry(value), 6)), .groups = "drop") |>
  unnest_wider(summary) |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```

## Evolvability (from G matrix) v2 
This is for a specific set of selection gradients:    
Beta_index = 1 is $\beta = \begin{bmatrix} 0 \\ 1 \end{bmatrix}$ = strong selection for longer tails only  
Beta_index = 2 is $\beta = \begin{bmatrix} -(1/\sqrt2) \\ -(1/\sqrt2) \end{bmatrix}$ = strong selection for shorter trunks and shorter tails  
Beta_index = 3 is $\beta = \begin{bmatrix} (1/\sqrt2) \\ -(1/\sqrt2) \end{bmatrix}$ = strong selection for longer trunks and shorter tails  
Multiply 'evolvability' by 100 to get in units of percent.  
```{r}
e.labels <- c(
  e = 'evolvability', r = 'respondability', c = 'conditional',
  a = 'autonomy (cB/eB)', i = 'integration'
)

e.params_beta |>
  pivot_longer(
    cols = -c(sample, Beta_index), 
    names_to = 'metric', 
    values_to = 'value'
  ) |> 
  mutate(
    metric = factor(e.labels[metric], levels = e.labels)
  ) |>  
  group_by(Beta_index, metric) |> 
  summarise(summary = list(round(vecSmry(value), 6)), .groups = "drop") |> 
  unnest_wider(summary) |> 
  arrange(metric, Beta_index) |> 
  pander::pander(split.tables = Inf, keep.line.breaks = TRUE)
```
