---
title: "Supplemental Materials - Model MCMC Diagnostic Plots"
output: pdf_document
knit:
  (function(inputFile, encoding) {
    rmarkdown::render(inputFile,
                      encoding = encoding,
                      output_dir = "Figures and Tables")
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  comment = NA,
  fig.height = 7,
  fig.width = 7
)

rm(list = ls())
library(tidyverse)
library(runjags)
options(scipen = 9999)
```

From the posterior of the model, we show histograms of the autocorrelation (AC.####), potential scale reduction factor (PSRF), and effective sample size (ESS) values for all parameters of interest in the model. Following these are the posterior predictive check distribution (PPD) expressed as the distribution of the deviation of the observed values from the PPD values. Finally, for the parameters of interest, we show the posterior trace, histogram, expected cumulative distribution (ECDF), and autocorrelation distribution.

```{r}
load('Model_outputs/posterior_20260605_0520.rdata')
```

### MCMC parameters  

Number of chains: `r length(post$mcmc)`  
Burnin: `r post$burnin`  
Samples per chain: `r post$sample`  
Total samples: `r post$sample * length(post$mcmc)`  
  
\newpage  

### A) Posterior distribution diagnostic summary
```{r}
post.smry$post |> 
  ggplot() +
  geom_histogram(aes(values), bins = 20) +
  facet_wrap(~diag, scales = 'free_x')
```

### B) Posterior predictive check distribution
```{r}
ggplot(ppc) +
  geom_histogram(aes(mean.diff), bins = 50) +
  facet_wrap(~ metric, scales = 'free_x') +
  labs(
    x = 'Length Difference (Observed - PPD)', 
    y = 'Count'
  )
```

\newpage
### C) By-parameter posterior summaries
```{r}
suppressMessages(plot(
  post, 
  vars = c(
    'deviance', 'sire.vcov', 'dam.vcov', 'int.vcov', 'resid.vcov'
  )
))
```